---
title: "R Notebook"
output: html_notebook
---

# Importing required libraries
```{r}

  library(tidyverse)
  library(readxl)

```
# Importing the data
```{r}
# The data downloaded from kaggle was in Excel format so the readxl 
# library had to be used to import the data.

    playerStatsLaLiga2019_2020 <- read_excel("La liga Player stats.xlsx")
```
# Questions

## 1. Is the Age of soccer players in the first division of the Spanish league(La Liga) normally distributed?

```{r}

ggplot(data = playerStatsLaLiga2019_2020) + 
  geom_histogram(aes(x = Age), binwidth=2, color="darkblue", fill="lightblue") + 
  scale_x_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35,40)) +
  geom_vline(aes(xintercept=mean(Age, na.rm = T)),
            color="blue", linetype="dashed", size=1) + 
  labs( title = "Distribution of soccer players ages", 
        subtitle = "Players from La Liga season 2019-2020", 
        caption = "data gathered from: kaggle.com") + 
  ylab("Number of players")

```

```{r}
# Display all the measurements of center plus standard deviation and range.
  summary(playerStatsLaLiga2019_2020$Age)
  message("Standard Deviation: ", sd(playerStatsLaLiga2019_2020$Age, na.rm =T))
  message("Range: ", max(playerStatsLaLiga2019_2020$Age, na.rm =T) -
            min(playerStatsLaLiga2019_2020$Age, na.rm =T))

```

## 2. Are goals scored evenly distributed among Foward players in the first division of the Spanish league?

```{r}
#First we need to get the forwards by themselves
forwardPlayersOnly <- subset(playerStatsLaLiga2019_2020, Pos == "FW", Gls...11 > 0)


ggplot(data = forwardPlayersOnly) + 
  geom_histogram(aes(x = Gls...11), binwidth=3, color="darkblue", fill="cyan3")  +
  geom_vline(aes(xintercept=mean(Gls...11, na.rm = T)),
            color="blue", linetype="dashed", size=1) + 
  labs( title = "Distribution of strikers goals", 
        subtitle = "Strikers from La Liga season 2019-2020", 
        caption = "data gathered from: kaggle.com") + 
  ylab("Number of players") + xlab("Goals Scored")

```

```{r}
# Display all the measurements of center plus standard deviation and range.

  summary(forwardPlayersOnly$Gls...11)
  message("Standard Deviation: ", sd(forwardPlayersOnly$Gls...11, na.rm =T))
  message("Range: ", max(forwardPlayersOnly$Gls...11, na.rm =T) -
            min(forwardPlayersOnly$Gls...11, na.rm =T))

```

## 3. Do all players play the same amount of games, meaning that the matches played per player is a normal distribution.

```{r}

ggplot(data = playerStatsLaLiga2019_2020) + 
  geom_histogram(aes(x = MP), binwidth=2, color="black", fill="lightgreen")  +
  geom_vline(aes(xintercept=mean(MP, na.rm = T)),
            color="blue", linetype="dashed", size=1) + 
  labs( title = "Distribution of matches played per player", 
        subtitle = "Players from La Liga season 2019-2020", 
        caption = "data gathered from: kaggle.com") + 
  ylab("Number of players") + xlab("Matches Played")

```

```{r}
# Display all the measurements of center plus standard deviation and range.

  summary(playerStatsLaLiga2019_2020$MP)
  message("Standard Deviation: ", sd(playerStatsLaLiga2019_2020$MP, na.rm =T))
  message("Range: ", max(playerStatsLaLiga2019_2020$MP, na.rm =T) -
            min(playerStatsLaLiga2019_2020$MP, na.rm =T))

```

## 4. Is there disparity in playing time among Barcelon's (The best soccer team in the world) soccer players?

```{r}
#Subsetting the Barcelona's soccer players
bestTeamOnly <- subset(playerStatsLaLiga2019_2020, Squad == "Barcelona")

ggplot(data = bestTeamOnly) + 
  geom_histogram(aes(x = Min), binwidth=275 , color="darkorange3", fill="darkorange")  +
  geom_vline(aes(xintercept=mean(Min, na.rm = T)),
            color="blue", linetype="dashed", size=1) + 
  labs( title = "Distribution of minutes played by Barcelona's players", 
        subtitle = "Barcelona F.C. players from the season of 2019-2020", 
        caption = "Barcelona F.C. is the best soccer team that ever existed") + 
  ylab("Number of players") + xlab("Minutes Played")

```

```{r}
# Display all the measurements of center plus standard deviation and range.

  summary(bestTeamOnly$Min)
  message("Standard Deviation: ", sd(bestTeamOnly$Min, na.rm =T))
  message("Range: ", max(bestTeamOnly$Min, na.rm =T) -
            min(bestTeamOnly$Min, na.rm =T))

```

